Smart Spreading Factor Assignment for LoRaWANs

### 2019-02-26 paper.tex
extend abbreviations

\section{LoRa} \label{LoRa}
\par LoRa is a proprietary physical layer radio/chipset technology that provides wireless link solution for LPWANs. It uses proprietary spread spectrum modulation technique that is the derivative of Chirp Spread Spectrum (CSS). A chirp is a sinusoidal signal of which frequency increases over time. Chirp frequency increases linearly and sweeps the entire bandwidth \cite{AN1200.22}.

@@

Where, $R_{b}$ is data rate in bps, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.

@@

\subsection{Spreading Factor Assignment Issue}
\par Simultaneous different SF transmissions are orthogonal to each other up to some extent. Which means that, a LoRa gateway (GW) can simultaneously receive multiple transmissions with different SFs. However, simultaneous transmissions with the same SF may not be received by the GW due to collision. For this reason, SF assignment of nodes is crucial for network performance.

\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a GW. However, a node can guess the distance from a GW by observing received signal power of a downlink transmission. If received signal power of a downlink transmission is too high, then node can decrease its next transmission SF to decrease power consumption. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper. Lowest possible SF assignment scheme is commonly used in LoRaWAN deployments. Also, a GW can request from a node to decrease its SF or transmit power.

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. The end devices close to the GW will fall into lowest SF (SF7) area section. The end devices close to the GW will probably select the lowest SF most of the time. Hence the number of collisions will increase as the number of end devices close to the GW increases.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision}
\caption{Collision between nodes close to the GW.}
\label{fig:collision}
\end{figure}

\section{LoRaWAN} \label{LoRaWAN}
\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions. LoRaWAN is developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN. LoRaWAN is based on pure ALOHA medium access which means that end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision. A typical LoRaWAN network consists of following three network entities.

\subsection{End node}
\par LoRaWAN end node (EN) is a low power embedded device that only communicates to GWs. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C. Different classes provide LPWAN solutions to different applications and deployments. Class A end nodes generate uplink transmission at any time and only receive a period of time after uplink transmission. Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission. Receive window is synchronized using a beacon packet transmitted by GWs. Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission. This provides Class C end nodes with low latency downlink communication, which requires more power consumption. In this paper, only Class A end devices are considered since Class A behavior leads to the lowest power consumption.

\subsection{Gateway}
\par LoRaWAN gateway (GW) is a device that communicates with the end nodes. A typical GW can receive from multiple channels at the same time. GWs are usually connected to power grid, so power consumption of a GW is insignificant in most of the deployments.

@@

\par In \cite{8090518}, another LoRaWAN ns-3 simulator is presented. Authors introduced an error model for determining range as well as interference between multiple simultaneous LoRa transmissions. Their simulator supports LoRaWAN Class A end devices, multiple GWs, both upstream and downstream confirmed messages. Their results show that allocating network parameters such as SF is highly important for the performance of LoRaWAN networks.

\par In \cite{8267219} and \cite{8430542}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal. Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.

@@

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even though they are able to communicate with lower SFs. This has potential to prevent collisions due to the orthogonality of the SFs as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, distribution of SF among nodes becomes an important problem. Increasing a node's SF should be done carefully since higher SF means longer air time and longer air time means increasing the probability of collisions with other high SF transmissions. In multiple GWs scenarios, this approach may increase the collisions with the nodes in other GW's range. Thus, extra care should be taken for nodes in the intersection area of the GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules. For this reason, machine learning based SF assignment approach is proposed to decrease the collisions for the same SF transmissions. This technique starts by learning the transmission behavior of the nodes in a network. An NS can keep track of successful uplink transmissions and their SFs. This NS can also keep track of some of the collided transmissions if header part of the packet is not interfered at the GW. However, NS cannot keep track of transmissions with lower receive power than sensitivity of the GW. Using this information, NS can train a classifier to predict future transmission result for a specific node and a specific SF. Using this prediction model NS can assign SFs to nodes considering the collision probability.

@@

\par It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator. Data set size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy. In real world deployments, NS can keep track of  transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.

\par To come up with a solution for SF assignment, DTC and SVM prediction schemes are integrated to the simulator. A classifier is trained from transmission logs and this classifier is used for selecting optimum SF for the nodes. Simulator first runs a simulation by assigning random SFs to the nodes. Simulation logs are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and the result of the transmission. Possible transmission results are successful, interfered or under sensitivity. These data are fed to Python scikit-learn DTC or SVM classifiers. After the classifier model is generated, a second simulation is run with classifier. Simulator selects optimum SF considering the prediction of transmission. If a transmission of a node with the lowest possible SF is predicted as interfered by the classifier, then simulator assigns higher SF and evaluates the new prediction. If no SF is predicted as successful, then simulator again assigns the lowest possible SF.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision_solution_single_gw}
\caption{Collision avoidance by using higher SF for nodes close to the GW.}
\label{fig:collision_solution_single_gw}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision_solution_multi_gw}
\caption{Collision avoidance for intersecting GWs.}
\label{fig:collision_solution_multi_gw}
\end{figure}

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study the effects of different SF strategies in LoRaWANs. LoRaWAN SF simulation tool source code is available at \cite{simlorasf}. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed circular shape network topology with input parameters; radius (m), number of nodes and number of GWs. Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and SF assignment method. With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J). Simulator also produces prediction accuracy and confusion matrix for machine learning schemes.

\par The simulator only covers the LoRaWAN Class A devices. Transmissions are always initiated by end nodes in pure ALOHA manner. Nodes generate a new packet randomly according to a Poisson interval for given packet rate parameter. Downlink transmissions are not considered. Downlink transmissions are rare in real world deployments since ISM band regulations dictate duty cycle limit for all devices including GWs.

@@

\par Receive sensitivity of a LoRa GW for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.

@@

\par If the received signal power is higher than the GW sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission.

@@

\section{Simulation Results} \label{Simulation Results}
\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 sec, packet generation rate = 0.01 pac/sec.

\par In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown. Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1. Increasing SFs increases air time. It increases the number of collisions thus decreases the PDR. High SF schemes gives poor PDR results as the number of nodes increases. Since network topology radius is quite small, all SFs can reach to GW.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{sf_pdr_r3000_g1_p0.01_s3600}.png}
\caption{PDR for various SFs.}
\label{fig:sf_pdr}
\end{figure}

\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of GWs is shown. Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used. Increasing number of GWs, decreases the SFs of nodes and decreases air time. This decreases number of collisions, hence, increases the PDR of network when network radius is constant.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{gw_pdr_r3000_p0.01_s3600}.png}
\caption{PDR for various number of GWs.}
\label{fig:gw_pdr}
\end{figure}

\par In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown. Number of GW is set to 1 and lowest SF assignment scheme is used. Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{r_pdr_g1_p0.01_s3600}.png}
\caption{PDR for various network radii.}
\label{fig:r_pdr}
\end{figure}

\par PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}. Randomly generated network radius is set to 5000 meters and number of GWs is set to 3. Prediction model needs nodes' locations and three GWs are enough to locate nodes positions by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases. Increasing number of nodes, increases number of interferences. Smart schemes improve network performance when LoRa interference is high. Moreover, smart schemes give better results when nodes are deployed closer to the GW, since nodes have margin to increase their SFs when they are deployed closer to the GW. If a node is far away from GW, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.

\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy. Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{prediction_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR for lowest SF and smart SF schemes.}
\label{fig:prediction_pdr}
\end{figure}

\begin{table}
\centering
\caption{PDR for lowest SF and smart SF schemes.}
\label{table:prediction_pdr}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                            & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                             & 100           & 500           & 1000          \\ \hline
\multirow{12}{*}{\textbf{r (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
                                 &                        & SVM    & 98.0          & 88.2          & 75.2          \\ \cline{3-6}
                                 &                        & DTC    & 97.8          & 89.8          & 78.7          \\ \cline{2-6}

@@

\begin{table}
\centering
\caption{Prediction accuracy for SVM and DTC.}
\label{table:prediction_accuracy}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
\multirow{8}{*}{\textbf{r (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}

@@

\par For space constraints, we omit results like network throughput and total transmit energy consumption. We invite readers to experiment the simulation tool with the parameters they like \cite{simlorasf}.

\section{Conclusion} \label{Conclusion}
\par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.

\par As for future work, transmit power optimization can be included to the proposed smart schemes. In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to GWs can decrease transmit power to save energy. This will make transmissions more vulnerable to interference thus requires extra care. Moreover, other machine learning methods can be investigated for SF assignment enhancement. Reinforcement learning could be a good candidate. Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve prediction performance.

>>>>

\section{LoRa} \label{LoRa}
\par LoRa is a proprietary physical layer radio/chipset technology that provides wireless link solution for low power wide area networks. It uses proprietary spread spectrum modulation technique that is the derivative of Chirp Spread Spectrum (CSS). A chirp is a sinusoidal signal of which frequency increases over time. Chirp frequency increases linearly and sweeps the entire bandwidth \cite{AN1200.22}.

@@

Where, $R_{b}$ is data rate in bit/sec, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.

@@

\subsection{Spreading Factor Assignment Issue}
\par Simultaneous different SF transmissions are orthogonal to each other up to some extent. Which means that, a LoRa gateway can simultaneously receive multiple transmissions with different SFs. However, simultaneous transmissions with the same SF may not be received by the gateway due to collision. For this reason, SF assignment of nodes is crucial for network performance.

\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a gateway. However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission. If received signal power of a downlink transmission is too high, then node can decrease its next transmission SF to decrease power consumption. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper. Lowest possible SF assignment scheme is commonly used in LoRaWAN deployments. Also, a gateway can request from a node to decrease its SF or transmit power.

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single gateway is illustrated. Different color rings represent achievable range of different SFs from the gateway and different color circles represent selected SF of the nodes. The end devices close to the gateway will fall into lowest SF (SF7) area section. The end devices close to the gateway will probably select the lowest SF most of the time. This causes a lot of collisions between same SF transmissions. Hence the number of collisions will increase as the number of end devices close to the gateway increases.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision}
\caption{Collision between nodes close to the gateway.}
\label{fig:collision}
\end{figure}

\section{LoRaWAN} \label{LoRaWAN}
\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions. LoRaWAN is developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN. LoRaWAN provides inter-operability between different LoRa networks. LoRa can be used as a wireless link technology without complying LoRaWAN, however this would break inter-operability between different LoRa networks. LoRaWAN is based on pure ALOHA medium access which means that end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision. A typical LoRaWAN network consists of following three network entities.

\subsection{End node}
\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C. Different classes provide LPWAN solutions to different applications and deployments. Class A end nodes generate uplink transmission at any time and only receive a period of time after uplink transmission. Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission. Receive window is synchronized using a beacon packet transmitted by gateways. Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission. This provides Class C end nodes with low latency downlink communication, which requires more power consumption. In this paper, only Class A end devices are considered since Class A behavior leads to the lowest power consumption.

\subsection{Gateway}
\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes. A typical gateway can receive from multiple channels at the same time. gateways are usually connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.

@@

\par In \cite{8090518}, another LoRaWAN ns-3 simulator is presented. Authors introduced an error model for determining range as well as interference between multiple simultaneous LoRa transmissions. Their simulator supports LoRaWAN Class A end devices, multiple gateways, both upstream and downstream confirmed messages. Their results show that allocating network parameters such as SF is highly important for the performance of LoRaWAN networks.

\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal. Their experimental results show that this power difference is around 16 dB. Such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.

\par In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs. They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the gateway for modeling interference in multiple gateways scenarios. Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.

@@

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even though they are able to communicate with lower SFs. This has potential to prevent collisions due to the orthogonality of the SFs as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, distribution of SF among nodes becomes an important problem. Increasing a node's SF should be done carefully since higher SF means longer air time and longer air time means increasing the probability of collisions with other high SF transmissions. In multiple gateways scenarios, this approach may increase the collisions with the nodes in other gateway's range. Thus, extra care should be taken for nodes in the intersection area of the gateways illustrated in Figure \ref{fig:collision_solution_multi_gw}.

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules. For this reason, machine learning based SF assignment approach is proposed to decrease the collisions for the same SF transmissions. This technique starts by learning the transmission behavior of the nodes in a network. An NS can keep track of successful uplink transmissions and their SFs. This NS can also keep track of some of the collided transmissions if header part of the packet is not interfered at the gateway. However, NS cannot keep track of transmissions with lower receive power than sensitivity of the gateway. Using this information, NS can train a classifier to predict future transmission result for a specific node and a specific SF. Using this prediction model NS can assign SFs to nodes considering the collision probability.

@@

\par It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator. Data set size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy. In real world deployments, NS can keep track of  transmissions and it can create a classifier daily basis, then gateways can request from nodes to use suggested SFs.

\par To come up with a solution for SF assignment, DTC and SVM prediction schemes are integrated to the simulator. A classifier is trained from transmission logs and this classifier is used for selecting optimum SF for the nodes. Simulator first runs a simulation by assigning random SFs to the nodes. Simulation logs are combined into four data columns which are: X and Y coordinates of the transmission source, SF of the transmission and the result of the transmission. Possible transmission results are successful, interfered or under sensitivity. These data are fed to Python scikit-learn DTC or SVM classifiers. After the classifier model is generated, a second simulation is run with classifier. Simulator selects optimum SF considering the prediction of transmission. If a transmission of a node with the lowest possible SF is predicted as interfered by the classifier, then simulator assigns higher SF and evaluates the new prediction. If no SF is predicted as successful, then simulator again assigns the lowest possible SF.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision_solution_single_gw}
\caption{Collision avoidance by using higher SF for nodes close to the gateway.}
\label{fig:collision_solution_single_gw}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision_solution_multi_gw}
\caption{Collision avoidance for intersecting gateways.}
\label{fig:collision_solution_multi_gw}
\end{figure}

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study the effects of different SF strategies in LoRaWANs. LoRaWAN SF simulation tool source code is available at \cite{simlorasf}. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed circular shape network topology with following input parameters: radius in meter (m), number of nodes and number of gateways. Global simulation input parameters are simulation duration in second (s), packet size in Byte (B), packet generation rate in packets per second (pac/sec) and SF assignment method. With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput in bits per second (bit/sec) and total transmit energy consumption in Joule (J). Simulator also produces prediction accuracy and confusion matrix for machine learning schemes.

\par The simulator only covers the LoRaWAN Class A devices. Transmissions are always initiated by end nodes in pure ALOHA manner. Nodes generate a new packet randomly according to a Poisson interval for given packet rate parameter. Downlink transmissions are not considered. Downlink transmissions are rare in real world deployments since ISM band regulations dictate duty cycle transmission limit for all devices including gateways.

@@

\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.

@@

\par If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission.

@@

\section{Simulation Results} \label{Simulation Results}
\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds, packet generation rate = 0.01 pac/sec.

\par In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown. Randomly generated network topology radius is set to 3000 meters and number of gateways is set to 1. Increasing SFs increases air time. It increases the number of collisions thus decreases the PDR. High SF schemes gives poor PDR results as the number of nodes increases. Since network topology radius is quite small, all SFs can reach to gateway.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{sf_pdr_r3000_g1_p0.01_s3600}.png}
\caption{PDR for various SFs. (r = 3000 m, GW = 1)}
\label{fig:sf_pdr}
\end{figure}

\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of gateways is shown. Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used. Increasing number of gateways, decreases the SFs of nodes and decreases air time. This decreases number of collisions, hence, increases the PDR of network when network radius is constant.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{gw_pdr_r3000_p0.01_s3600}.png}
\caption{PDR for various number of GWs. (r = 3000 m, SF = SF{\_}Lowest)}
\label{fig:gw_pdr}
\end{figure}

\par In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown. Number of gateway is set to 1 and lowest SF assignment scheme is used. Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{r_pdr_g1_p0.01_s3600}.png}
\caption{PDR for various network radii. (GW = 1, SF = SF{\_}Lowest)}
\label{fig:r_pdr}
\end{figure}

\par PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}. Randomly generated network radius is set to 5000 meters and number of gateways is set to 3. Prediction model needs nodes' locations and three gateways are enough to locate position of nodes by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases. Increasing number of nodes, increases number of interferences. Smart schemes improve network performance when LoRa interference is high. Moreover, smart schemes give better results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway. If a node is far away from gateway, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.

\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy. Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high. LoRaWAN simulation usually produces imbalanced dataset, DTC gives better accuracy for interfered transmission labels.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{prediction_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR for lowest and smart SF schemes. (r = 5000 m, GW = 3)}
\label{fig:prediction_pdr}
\end{figure}

\begin{table}
\centering
\caption{PDR for lowest and smart SF schemes. (GW = 3)}
\label{table:prediction_pdr}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                            & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                             & 100           & 500           & 1000          \\ \hline
\multirow{12}{*}{\textbf{r (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
                                 &                        & SVM    & 98.0          & 88.2          & 75.2          \\ \cline{3-6}
                                 &                        & DTC    & 97.8          & 89.8          & 78.7          \\ \cline{2-6}

@@

\begin{table}
\centering
\caption{Prediction accuracy for SVM and DTC. (GW = 3)}
\label{table:prediction_accuracy}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
\multirow{8}{*}{\textbf{r (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}

@@

\par For space constraints, we omit simulation results like network throughput, total transmit energy consumption and confusion matrices. We invite readers to experiment the simulation tool with the parameters they like \cite{simlorasf}.

\section{Conclusion} \label{Conclusion}
\par In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed. We present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to gateways.

\par As for future work, transmit power optimization can be included to the proposed smart schemes. In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to gateways can decrease transmit power to save energy. This will make transmissions more vulnerable to interference thus requires extra care. Moreover, other machine learning methods can be investigated for SF assignment enhancement. Reinforcement learning could be a good candidate. Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve prediction performance.
